Variation on a Variational Principle

### 2020-09-23 draft1.tex
added lemma, TODO finish proof

@@

We obtain an angle-splitting by applying the
Feasible Flow theorem TODO as follows:
Consider the directed graph whose vertex set is
$E \cup F \cup \{\otimes\}$,
where $E, F$ are the set of edges, faces in $\Gamma_T(K)$,
and $\otimes$ is some abstract vertex.
There is a directed edge
\begin{itemize}
\item $\otimes \to f$ for each face $f\in F$, with capacity interval
	$[\pi, \infty)$,
\item $f \to e$ for each edge $e \in \del f$,
	with capacity interval $[\veps, \infty)$
	for some $\veps>0$ to be set later,
\item $e \to \otimes$ for each edge $e$, with capacity interval
	$(-\infty, \pi/2]$.
\end{itemize}
By TODO lemma about $2|F'| \leq |E'|$,
and taking $\veps < \pi / |\text{max face size}|$,
the feasible flow condition is satisfied,
so a feasible flow exists.
Since $2|F| = |E|$, the capacity interval restrictions
on the flow at $\otimes$ is sharp, so out-edges at $\otimes$
have flow $\pi$ and in-edges at $\otimes$ have flow $\pi/2$.
Then the flow $f \to e$ gives us $\vphi_{\vec{e}}$,
where $f$ is the face to the left of $\vec{e}$.
(we adapted this argument from \cite{BandS}).

>>>>

\begin{lemma}
\label{l:2fe}
Let $G$ be a 4-valent graph on $\torus$ with no bigons or self-loops,
and whose faces can be checkerboard colored.
Then for any subset $F' \subseteq F$ of faces,
\[
	2|F'| \leq |E'|
\]
where $E' = \cup_{f \in F'} \del f$ is the set of edges
that meets some face of $F'$.
\end{lemma}

\begin{proof}
We induct on $|F'|$ (over all possible such graphs $G$);
for $F' = \emptyset$, there is nothing to prove.

Call a vertex or edge \emph{interior} if it does not meet
$F'' = F \backslash F'$.
Suppose there is an interior vertex $v$.
Make the modification as in \figref{TODO}.
This decreases $|F'|$ by one and $|E'|$ by 2.
However, this process may produce bigons;
if it does, perform the modification as in \figref{TODO}
simultaneously on all bigons,
repeatedly until there are no more bigons.
%each bigon eliminated decreases $|F'|$ by one and $|E'|$ by 2,

TODO finish proof..

We must check that this process terminates with a graph
which satisfies the hypotheses set out in the lemma.
It is clear that checkboard colorability is maintained,
obviously there will not be any bigons.
After the modification of the internal vertex,
at most two bigons are created;
in any case, the bigons belong to one color of the 
It is clear that when a bigon is eliminated,
the only faces which it is adjacent to will have a reduced number of edges.
\end{proof}

@@

We obtain an angle-splitting by applying the
Feasible Flow theorem TODO as follows:
Consider the directed graph whose vertex set is
$E \cup F \cup \{\otimes\}$,
where $E, F$ are the set of edges, faces in $\Gamma_T(K)$,
and $\otimes$ is some abstract vertex.
There is a directed edge
\begin{itemize}
\item $\otimes \to f$ for each face $f\in F$, with capacity interval
	$[\pi, \infty)$,
\item $f \to e$ for each edge $e \in \del f$,
	with capacity interval $[\veps, \infty)$
	for some $\veps>0$ to be set later,
\item $e \to \otimes$ for each edge $e$, with capacity interval
	$(-\infty, \pi/2]$.
\end{itemize}
By \lemref{l:2fe}about $2|F'| \leq |E'|$,
and taking $\veps < \pi / |\text{max face size}|$,
the feasible flow condition is satisfied,
so a feasible flow exists.
Since $2|F| = |E|$, the capacity interval restrictions
on the flow at $\otimes$ is sharp, so out-edges at $\otimes$
have flow $\pi$ and in-edges at $\otimes$ have flow $\pi/2$.
Then the flow $f \to e$ gives us $\vphi_{\vec{e}}$,
where $f$ is the face to the left of $\vec{e}$.
(we adapted this argument from \cite{BandS}).
